% !TeX root = ./0Base.tex

\chapter{Introduction}

Complex web applications keep becoming more popular in recent years. They are very easily accessible from any place in the world and can run on almost any modern device, as it requires only web browser and the Internet connection to run.

Web frameworks simplify the development process of web applications significantly improving developers productivity. There is a huge variety of choices between the mentioned software, so choosing one may be a difficult task.

The choice of server-side framework is crucial, as it is responsible for handling sensitive data and plays a key role in the overall performance of the application.

The goal of this document is to focus on three of the most popular server-side frameworks and compare their performance under high load as well as basic security measurements. The results of this thesis will be important for web developers and architects, that need to decide on which framework should they choose for their application.

Web frameworks for this comparison were chosen from the list of Stack Overflow Developer Survey 2020 Web Frameworks popularity \cite{devSurveyWeb}. The chosen frameworks (and their respective languages) are:
\begin{itemize}
    \item Express.js (JavaScript)
    \item ASP.NET (C\#)
    \item Django (Python)
\end{itemize}

Chapter 2 describes chosen frameworks, including their architecture and requirements. For the definition of the criteria against which the results will be measured, see Chapter 3. Chapter 4 presents design of the system - database models, application structure and environment preparation. Chapter 5 shows framework specific application implementation details. Results of the tests and comparison of the technologies can be found in chapter 6. For final conclusion of this document, see chapter 7.
